Transition path sampling based calculation of free energies of enzymatic reactions

### 2022-04-07 tex/manuscript.tex
added some more figures and text

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]
  {Transition path sampling and the calculation of free energies of enzymatic
  reactions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{TPS,NMR,UV}
\keywords{American Chemical Society, \LaTeX}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimations of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations are not employed in these cases because 
trajectories that take the system from reactant state to product state through
a transition state are rare and would take long time

Ergodic sampling of the phase space within straighforward MD simulations 
is often difficult to achieve with the presence of energy barriers. 
To overcome such barriers enhanced sampling techniques 
have been developed over the years such as 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} etc. These methods rely on applying bias
potentials for the system to overcome energy barriers and hence 
require the knowledge of the reaction coordinates.  
Transition path sampling \cite{Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}
(TPS) is an enhanced sampling method for 
rare events that transition a system between two stable states often 
referred to as reactants and products.

TPS focuses on transition paths and not on stable states which introduces 
biases in the sampling of equilibrium distribution of the reaction coordinate. \cite{Dellago09AdvCompSimAppp167}

TPS based free energy simulations have been attempted previously by 
Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111}

@@

%------------------------------------------------------------------------------ 
\subsection{Adenosine deaminase reaction coordinates and free energy}
%------------------------------------------------------------------------------ 
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.22]{./figures/ada-protein.png}
\end{figure}

@@

%\subsection{Outline}
\begin{figure}[ht!]
\centering
\begin{minipage}[b]{0.45\linewidth}
\includegraphics[scale=0.4]{figures/reac-sc-oc-167.png}
\caption{Trajectory 167}
\label{fig:minipage1}
\end{minipage}
\quad
\begin{minipage}[b]{0.45\linewidth}
\includegraphics[scale=0.8]{figures/den-fenergy-1.pdf}
\caption{Free energy profile for the enzymatic reaction in human Mat2a enzyme}
\label{fig:minipage2}
\end{minipage}
\end{figure}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% The document title should be given as usual. Some journals require
%% a running title from the author: this should be supplied as an
%% optional argument to \title.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\title[]
  {Transition path sampling based calculation of free energies of enzymatic
  reactions}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Some journals require a list of abbreviations or keywords to be
%% supplied. These should be set up here, and will be printed after
%% the title and author information, if needed.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\abbreviations{TPS,NMR,UV}
\keywords{American Chemical Society, \LaTeX}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%% Start the main part of the manuscript here.
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
Molecular dynamics (MD) simulations are increasingly being
used to calculate quantitative estimations of experimentally measurable 
kinetic and thermodynamic parameters of enzyme catalyzed reactions. 
\cite{Karplus02NatStructMolBiol9p646,
Schramm98AnnuRevBiochem67p693,Zhang05AccChemRes38p379,
Schramm11AnnuRevBiochem80p703,Schramm18ChemRev118p11194} 
This has helped in proposing mechanisms and explanations for 
enzyme activity, fast high throughput screening of thousands of inhibitor 
molecues in drug lead discovery, \cite{Jorgensen09AccChemRes42p724,Sliwoski14PharmacolRev66p334} 
directed evolution for designing enzymes with enhanced catalytic 
activities, \cite{Thyme09Nature461p1300,Bloom09PNAS106p9995,Schafer19JAmChemSoc141p10431} etc.    
Straightforward MD simulations adequately sample 
the long-lived stable reactant and product states 
but the high energy barrier-crossing events that occur rarely and at long 
time scales are not easily accessible using this method alone.
To access such states at a higher frequency using MD simulations 
several enhanced sampling techniques 
have been developed in the past few decades some of which are 
umbrella sampling, \cite{Kastner11WileyInterdiscipRevComputMolSci1p932} 
metadynamics, \cite{Barducci11WileyInterdiscipRevComputMolSci1p826}
steered MD, \cite{Park04JChemPhys120p5946}
milestoning, \cite{Faradjian04JChemPhys120p10880} and transition path sampling 
(TPS). \cite{Pratt86JChemPhys85p5045,Bolhuis02AnnRevPhysChem53p291,Dellago98JChemPhys108p1964}

TPS in particular is an attractive method since it samples reactive trajectories
in which the rare barrier-crossing events are guaranteed to occur and
it does not require a priori knowledge of the reaction coordinate which can be 
a complicated entity in large biological systems. TPS has been successfully 
applied to study problems such as enzyme catalysis, protein folding, etc. 
The TPS ensemble which is used to obtain a unique statistical definition for the reaction coordinate
and transition state of a reaction consists of so called reactive trajectories which 
link the stable reactant and product states. In this Monte-Carlo procedure in the trajecotry space
rejected moves correspond to non-reactive trajectories. But these trajectories are 
required to obtain the equilibrium distribution function.

The TPS ensemble consists of reactive trajectories which connect a reactant state to
a product state by going through the transition state and the probabilities of configurations thus sampled 
do not replicate the equilibrium probabilities. To calculate equilibrium properties such as the free energy 
of a reaction, equilibrium sampling of the configurations along a reaction coordinate is 
required which is not obtained with straightforward TPS. \cite{Dellago09AdvCompSimAppp167}

TPS based free energy simulations have been attempted previously by 
Brotzakis et al., \cite{Brotzakis19JChemPhys151p174111,Dellago04MolSimul30p795}

@@

%------------------------------------------------------------------------------ 
\subsection{Human MethionineAdenosyltransferase 2A}
%------------------------------------------------------------------------------ 
The SN2 reaction catalyzed by the human Mat2a enzyme leads to the production of 
S-Adenosylmethionine compound which is part of the ATP cycle \cite{Niland21Biochem60p791}
\begin{figure}[ht!]
\includegraphics[scale=0.5]{figures/comm-80-mat2a.pdf}
\end{figure}

\begin{figure}
\includegraphics[scale=0.5]{figures/mat2a-fenergy.pdf}
\caption{The free energy as a function of the order parameter.}
\end{figure}
%------------------------------------------------------------------------------ 
\subsection{Adenosine deaminase reaction coordinates and free energy}
%------------------------------------------------------------------------------ 
\begin{figure}[ht!]
\centering
\includegraphics[scale=0.22]{./figures/ada-protein.png}
\end{figure}

@@

%\subsection{Outline}
%\begin{figure}[ht!]
%\centering
%\begin{minipage}[b]{0.45\linewidth}
%\includegraphics[scale=0.4]{figures/reac-sc-oc-167.png}
%\caption{Trajectory 167}
%\label{fig:minipage1}
%\end{minipage}
%\quad
%\begin{minipage}[b]{0.45\linewidth}
%\includegraphics[scale=0.8]{figures/den-fenergy-1.pdf}
%\caption{Free energy profile for the enzymatic reaction in human Mat2a enzyme}
%\label{fig:minipage2}
%\end{minipage}
%\end{figure}
